\plaintitle

### 2017-01-08 drafts/labor/complexity.tex
complexity section revisions

\subsection[What are the complexity limits of on--demand work]
{Identifying the Complexity Limits of On--Demand Work}\label{sec:complexity}
A key question to the future of on--demand work is
\textit{what} precisely will become part of this economy.
Paid crowdsourcing began with simple microtasks on platforms such as
Amazon Mechanical Turk, but
microtasks are only helpful if they build up to a larger whole.
So, our first question:
how complex can the work outcomes from on--demand work be?

@@

\topic{Crowdsourcing research has spent the better part of a decade
proving the viability of crowdsourcing in complex work.}
Unless on--demand work as a whole can demonstrate viability for meaningfully complex tasks,
the argument runs,
it will be incapable of ensuring a pro--social outcome for work and workers~\cite{crowdworkFuture}.
\citeauthor{crowdForgeKittur} first opened the question of
whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.
Their work demonstrated proof--of--concept crowdsourcing of
a simple encyclopedia article and news summary
--- tasks which could be verified or repeated
with reasonable expectations of similar outcomes.
Seeking to raise the complexity ceiling~\cite{myers2000past},
researchers have since created
additional proof--of--concept applications and techniques,
including conversational assistants~\cite{lasecki2013chorus},
medical data interpreters~\cite{lasecki2013chorus}, and
idea generation work--flows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
to name a few examples.

\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new crowdsourcing work--flows.}
Beginning with a goal that has
presented significant challenges for computers,
the researcher leverages an insight from Computer Science
(for example, MapReduce~\cite{crowdForgeKittur} or
sequence alignment algorithms~\cite{lasecki2012real})
and arranges humans as computational black boxes within those approaches.
This approach has proven a compelling one because
it leverages the in--built advantages of
scale,
automation, and
programmability that software affords.

\topic{It is now clear that this computational workflow approach works with focused complex tasks, but
the broader wicked problems largely remain unsolved~\cite{rittel1973dilemmas}.}
As a first example,
idea generation shows promise~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
but there is as yet no general crowdsourced solution for
the broader goal of invention and innovation~\cite{fuge2014analysis}.
Second,
focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,writingMicroTasks,agapie2015crowdsourcing}, but
there is no general solution to create
a cross--domain, high--quality crowd--powered author. 
Third,
data analysis tasks such as
clustering~\cite{chilton2013cascade},
categorization~\cite{andre2014crowd}, and
outlining~\cite{luther2015crowdlines}
are possible, but there is no general solution for sense--making.
It is not yet clear what insights would be required
to enable crowdsourced solutions for these broader wicked problems.

@@

\topic{Piecework's body of research most squarely addresses complexity in two of the cases we looked at earlier:
Airy's human computers and among industrial workers.}
The majority of piecework's perspective, on this topic, will come from these areas.

@@

However, when piecework intially entered the American economy,
it was not used for complex work.
Without having designed processes to combine work intelligently, piecework managers were restricted to available workers' skills: it was infeasible to provide new pieceworkers with the comprehensive education
that apprenticeships imparted \cite{hart2013rise}.
So, initially piecework arose for farm work, and as
\citeauthor{hughRaynbirdTaskWork} and others discuss,
the practice remained relatively obscure until
it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
Complexity levels remained low at the turn of
the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.

\topic{Measurement also limited the complexity of piecework: only tasks that could be measured and priced could be completed via piecework.}
Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers.
They identified task homogeneity and measurement as key requirements for piecework.
However, complex, creative work
--- which is inherently heterogeneous and difficult to routinize ---
would then be unsuitable~\cite{10.2307/23702539}.
\citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
``efficiency experts'' can effectively gauge how long known tasks should take, but
would find themselves overwhelmed if they attempted to assess creative tasks like scientific research,
which can take an arbitrary number of iterations before proceeding to a subsequent step.

@@

\topic{The research on piecework tells us that
we should expect piecework to thrive in industries where
the nature of the work is limited in complexity~\cite{Brown01041990}, and become less common as work becomes more complex.}
Given the flourishing of on--demand labor platforms such as
Uber, AMT, and others, we ask ourselves
what --- if anything --- has changed.

@@

\topic{Technology increases non--experts' levels of expertise by giving access to information that would otherwise be unavailable.}
For example, taxi drivers in London endure rigorous training to pass a test known as ``The Knowledge''
--- a demonstration of the driver's comprehensive familiarity with the city's roads.
This test is so challenging that veteran drivers develop significantly larger
the regions of the brain associated with spatial functions such as navigation~\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,skok2000managing,Woollett1407,woollett2011acquiring}.
In contrast, with on--demand platforms such as Uber, services such as Google Maps \& Waze make it possible for
people entirely unfamiliar with a city
to operate professionally~\cite{silva2013traffic,hind2014outsmarting}.
Other examples include search engines enabling information retrieval, and
word processors enabling spelling and grammar checking.
By augmenting the human intellect~\cite{engelbart2001augmenting},
computing has shifted the complexity of work that is possible without training.

\topic{Algorithms have automated some tasks that previously fell to management.}
Computational systems now %hire workers~\cite{turkitLittle,weld2010decision}, as well as direct their activities~\cite{uberAlgorithm}, and 
act as ``piecework clerks''~\cite{10.2307/23702539} to inspect, modify and combine work~\cite{turkopticon,takingAHITMcInnis}.
%In many cases, the intermediary function has been removed as well, leading workers to need to directly email requesters for clarification and feedback~\cite{martin2014being}.
These algorithms are less able than human managers to evaluate subjective work.
So, assessment remains a bottleneck.

@@

% \clearpage
\subsubsection{\implication}
Algorithms are undoubtably capable of shepherding more complex work than the linear processes available to Airy and Ford.
However, as work becomes more complex, it becomes increasingly difficult to codify a process to achieve it~\cite{Faraj2006a,edmondson2012teaming}.
So, while algorithms will increase the complexity ceiling beyond what was possible previously with piecework, it seems likely that this increase is not enough to achieve truly complex work.

However, technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of on--demand work outcomes.
Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
of human computation~\cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework~\cite{hart2013rise}.
To some extent, an argument can be made that
MOOCs and other online education resources
provide crowd workers with the resources that they need, but 
it remains to be seen whether that work will be appropriately valued, let alone
properly interpreted by task solicitors~\cite{aguaded2013mooc}.
If we can overcome this obstacle,
we might be able to empower more of these workers to do complex work such as engineering,
rather than doom them to ``uneducated'' match--girl reputations~\cite{10.2307/3827491}.
However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to on--demand work unless it makes common expertise more broadly available.

Evaluation remains as difficult for crowd work as it did for the efficiency experts. 
Reputation systems for crowdsourcing platforms remain notoriously inflated~\cite{Horton2015a}.
Creating communities of practice for peer assessment~\cite{crowdguilds} may offer a path forward, but many aspects of assessment remain subjective: whether a logo made for a client is fantastic or terrible may depend on taste.

So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes than it has today.
Improvements in workflows, measurement and evaluation have already been made, and it is not clear that the remaining challenges are readily solveable.
On--demand work will be far more broadly distributed than piecework historically was---reaching many more tasks and areas of expertise by virtue of the internet---but not necessarily far more complex than what piecework previously achieved.

>>>>

\subsection[What are the complexity limits of on--demand work]
{Complexity Limits of On--Demand Work}\label{sec:complexity}
A key question to the future of on--demand work is
\textit{what} precisely will become part of this economy.
Paid crowdsourcing began with simple microtasks on platforms such as
Amazon Mechanical Turk, but
microtasks are only helpful if they build up to a larger whole.
So, our first question:
how complex can the work outcomes from on--demand work be?

@@

%\topic{Crowdsourcing research has spent the better part of a decade
%proving the viability of crowdsourcing in complex work.}
%Unless on--demand work as a whole can demonstrate viability for meaningfully complex tasks,
%the argument runs,
%it will be incapable of ensuring a pro--social outcome for work and workers~\cite{crowdworkFuture}.
\topic{\citeauthor{crowdForgeKittur} first opened the question of
whether crowdsourcing could be used for goals that are not simple parallel tasks~\cite{crowdForgeKittur}.}
Their work demonstrated proof--of--concept crowdsourcing of
an encyclopedia article and news summary
--- tasks which could be verified or repeated
with reasonable expectations of similar outcomes.
Seeking to raise the complexity ceiling,
researchers have since created
additional proof--of--concept applications and techniques,
including conversational assistants~\cite{lasecki2013chorus},
medical data interpreters~\cite{lasecki2013chorus}, and
idea generation work--flows~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a}.

\topic{To achieve complex work, this body of research has often applied ideas from Computer Science to design new crowdsourcing work--flows.}
The system designer leverages techniques such as MapReduce~\cite{crowdForgeKittur} and
sequence alignment algorithms~\cite{lasecki2012real}, arranging humans as computational black boxes within.
This approach has proven a compelling one because
it leverages the in--built advantages of
scale,
automation, and
programmability that software affords.

\topic{It is now clear that this computational workflow approach works with some classes of complex tasks, but
the broader wicked problems largely remain unsolved~\cite{rittel1973dilemmas}.}
As a first example,
idea generation shows promise~\cite{YuEncouragingOutside,yu2014distributed,Yu2016a},
but there is as yet no general crowdsourced solution for
the broader goal of invention and innovation~\cite{fuge2014analysis}.
Second,
focused writing tasks are now feasible~\cite{Kim2017,bernsteinSoylent,Nebeling:2016:WCW:2858036.2858169,writingMicroTasks,agapie2015crowdsourcing}, but
there is no general solution to create
a cross--domain, high--quality crowd--powered author. 
Third,
data analysis tasks such as
clustering~\cite{chilton2013cascade},
categorization~\cite{andre2014crowd}, and
outlining~\cite{luther2015crowdlines}
are possible, but there is no general solution for sense--making.
It is not yet clear what insights would be required
to enable crowdsourced solutions for these broader wicked problems.

@@

\topic{Piecework's body of research most squarely addresses complexity in two of the cases we looked at earlier:
Airy's human computers and among industrial workers.}
%The majority of piecework's perspective, on this topic, will come from these areas.

@@

However, when piecework intially entered the American economy,
it was not used for complex work.
Without having designed complex work processes, piecework managers were restricted to available workers' skills such as sewing: it was infeasible to provide new pieceworkers with the comprehensive education
that apprenticeships imparted \cite{hart2013rise}.
So, initially piecework arose for farm work, and as
\citeauthor{hughRaynbirdTaskWork} and others discuss,
the practice remained relatively obscure until
it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
Complexity levels remained low at the turn of
the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.

\topic{Measurement also limited the complexity of piecework: only tasks that could be measured and priced could be completed via piecework.}
Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers.
They identified task homogeneity and measurement as key requirements for piecework.
However, complex, creative work
--- which is inherently heterogeneous and difficult to routinize ---
was unsuitable~\cite{10.2307/23702539}.
\citeauthor{Brown01041990}'s description of ``efficiency experts'' would corroborate this;
efficiency experts can effectively gauge how long known tasks should take, but
would find themselves overwhelmed if they attempted to assess creative tasks like scientific research,
which can take an arbitrary number of iterations before proceeding to a subsequent step.

@@

\topic{The research on piecework tells us that
we should expect piecework to thrive in industries where
the nature of the work is limited in complexity~\cite{Brown01041990}, and become less common as work becomes more complex.}
%Given the flourishing of on--demand labor platforms such as
%Uber, AMT, and others, we ask ourselves
%what --- if anything --- has changed.
However, has computation shifted piecework's previous limits of expertise, measurement, and evaluation?

@@

\topic{In some ways, yes: technology increases non--experts' levels of expertise by giving access to information that would otherwise be unavailable.}
For example, taxi drivers in London endure rigorous training to pass a test known as ``The Knowledge'': a demonstration of the driver's comprehensive familiarity with the city's roads.
This test is so challenging that veteran drivers develop significantly larger
the regions of the brain associated with spatial functions such as navigation~\cite{Maguire11042000,Maguire2894,Skok:1999:KML:299513.299625,skok2000managing,Woollett1407,woollett2011acquiring}.
In contrast, with on--demand platforms such as Uber, services such as Google Maps and Waze make it possible for
people entirely unfamiliar with a city
to operate professionally~\cite{silva2013traffic,hind2014outsmarting}.
Other examples include search engines enabling information retrieval, and
word processors enabling spelling and grammar checking.
By augmenting the human intellect~\cite{engelbart2001augmenting},
computing has shifted the complexity of work that is possible without training.

\topic{Algorithms have automated some tasks that previously fell to management.}
Computational systems now %hire workers~\cite{turkitLittle,weld2010decision}, as well as direct their activities~\cite{uberAlgorithm}, and 
act as ``piecework clerks''~\cite{10.2307/23702539} to inspect, modify and combine work~\cite{turkopticon,takingAHITMcInnis}.
%In many cases, the intermediary function has been removed as well, leading workers to need to directly email requesters for clarification and feedback~\cite{martin2014being}.
However, these algorithms are less able than human managers to evaluate subjective work.
%So, assessment remains a bottleneck.

@@

% \clearpage
\subsubsection{\implication}
Algorithms are undoubtably capable of shepherding more complex work than the linear processes available to Airy and Ford.
However, as work becomes more complex, it becomes increasingly difficult to codify a process to achieve it~\cite{Faraj2006a,edmondson2012teaming}.
So, while algorithms will increase the complexity ceiling beyond what was possible previously with piecework, there is a fundamental limit to how complex such work can become.

Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of on--demand work outcomes.
Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
of human computation~\cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework~\cite{hart2013rise}.
To some extent, an argument can be made that
MOOCs and other online education resources
provide crowd workers with the resources that they need, but 
it remains to be seen whether that work will be appropriately valued, let alone
properly interpreted by task solicitors~\cite{aguaded2013mooc}.
If we can overcome this obstacle,
we might be able to empower more of these workers to do complex work such as engineering,
rather than doom them to ``uneducated'' match--girl reputations~\cite{10.2307/3827491}.
However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to on--demand work unless it makes common expertise more broadly available.

Evaluation remains as difficult for crowd work as it did for the efficiency experts. 
Reputation systems for crowdsourcing platforms remain notoriously inflated~\cite{Horton2015a}.
Ultimately, many aspects of assessment remain subjective: whether a logo made for a client is fantastic or terrible may depend on taste.

So, in the case of complexity, the history of piecework does not yet offer compelling evidence that on--demand work will achieve far more complex outcomes than piecework did.
Improvements in workflows, measurement and evaluation have already been made, and it is not clear that the remaining challenges are readily solveable.
However, on--demand work will be far more broadly distributed than piecework historically was---reaching many more tasks and areas of expertise by virtue of the internet.
